---
title: "Statistical Inference Part 1"
author: "Ozge Tugrul Sonmez"
date: "12/07/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Simulation Exercise

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with **rexp(n, lambda)** where lambda is the rate parameter. The mean of exponential distribution is **1/lambda** and the standard deviation is also **1/lambda**. Also, the variance of exponential distribution is **(1/lambda)^2**. 
Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.


```{r}
library(ggplot2)
library(reshape2)

nosim<-1000
n<-40
lambda<-0.2
set.seed(123)

```

### 1-Show the sample mean and compare it to the theoretical mean of the distribution.

```{r}
#Theoretical mean of the distribution

theoretical_mean<-1/lambda

theoretical_mean

#Mean of the Sample Means
mean(apply(matrix(rexp(n=nosim*n,rate=0.2),nosim),1,mean))

means<- NULL
for (i in 1 :nosim) means = c(means, mean(rexp(n, lambda)))

sample_and_theoretical_means<-data.frame(sample=means,theoretical=c(rep(1/lambda,1000)))

means_grouped<-data.frame(sample=mean(means),theoretical=c(5))

means_type<-melt(means_grouped,variable.name="means.type",value.name = "value")

ggplot(sample_and_theoretical_means,aes(x=sample))+geom_histogram(aes(y=..count..),
      binwidth=0.2,colour="dark orange",fill="bisque")+
  labs(x="Sample Means",y="Frequency",title="Histogram of the Sample Means")+
  geom_vline(data=means_type,aes(xintercept = value, colour = means.type), size = 2, 
             linetype = "dashed")
 

```

### 2-Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

Theoretical variance of exponential distribution is the square of the theoretical mean 
of exponential distribution 

Sample size is not used in theoretical variance or standard deviation, so the formula is:
**(1/lambda)^2**

```{r}
#Theoretical variance of the exponential distribution (lambda 0.2)

theoretical_variance<-(1/lambda)^2

theoretical_variance

#Variance of the Sample Means

(sd(apply(matrix(rexp(n=nosim*n,rate=0.2),nosim),1,mean)))^2

```

The variance of the sample means is **((1/lambda)^2)/n**. Since the variance of sample mean is related to the sample size,as the sample size n increases, the variance of the sample means decreases.


### 3-Show that the distribution is approximately normal.

Compared to the 1000 random exponentials, sample mean distribution looks far more Gaussian than the original exponential distribution.

The central limit theorem states that the sampling distribution of the sample means approaches a normal distribution as the sample size gets larger. This fact holds especially true for sample size over 30. In this example, sample size n is 40.

```{r}

#Simulated 1000 random exponentially distributed variables

exponential<-rexp(n=nosim,rate=0.2)

ggplot(as.data.frame(exponential),aes(x=exponential))+geom_histogram(binwidth=2,colour="red",fill="salmon")+
  geom_vline(xintercept = 5,colour="red",size=2)+labs(x="Values",y="Frequency",
       title="Exponential Distribution")


#Simulated 1000 exponentially distributed sample means (Approximately normally distributed)

ggplot(data=as.data.frame(means),aes(x=means))+geom_histogram(aes(y=..density..),
  binwidth=0.2,colour="darkgreen",fill="white")+
  scale_x_continuous(breaks=2:8)+
  stat_function(fun=dnorm,args=list(mean=mean(means,na.rm = TRUE),
                                    sd=sd(means,na.rm=TRUE)),colour="red",size=1)+
  labs(x="Sample Means",y="Density",
       title="Simulation of Exponentially Distributed Sample Means")+
  geom_density(colour="dark orange")


```

It can be seen that the sampling distribution is approximately normally distributed. If the samle size increases the sampling distribution will look more normal distribution.

```{r}
qqnorm(means)

qqline(means,col="red")
```



